Missing links and the topological robustness of food webs

### 2023-12-18 manuscript/ms_C3_v3.Rmd
Abstract shortened

 1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., real false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occurring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that model predicted food webs usually have greater connectance than the observed ones. Thus, we expect that predicted food webs with fewer missing links are more robust than the observed ones. However, this expectation has never, to our knowledge, been tested, nor has the effect size, i.e. the difference in robustness of predicted and observed food web with respect to the difference in their connectance, been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links. We did this for 12 different food webs from a wide variety of ecosystems. We used three extinction scenarios: random, most connected first, and least connected first.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, which can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found that the food webs robustness increased by 0.52 units for the most connected species extinction scenario and by 0.04 units for the random species extinction scenarios. On the other hand, we saw no effect or very little effect, on average, of increased connectance for the least connected species extinction scenario for all except two of the 12 food webs.
  5) These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Moreover, differences in the structural properties of the model predicted food webs, such as proportion of basal species and maximum trophich height can have large effects on robustness and therefore could be used to further constrain predicted food webs and perhaps produce more constrained predictions of food web robustness.

>>>>

 1)  Undersampling can lead to missing trophic interactions in recorded food webs, impacting the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., real false positives. 
  2) Previous research shows that (i) food web robustness increases with connectance, and (ii) that model predicted food webs usually have greater connectance than the observed ones. Thus, we expect that predicted food webs with fewer missing links are more robust than the observed ones. However, this expectation has never, to our knowledge, been tested, nor has the effect size, i.e. the difference in robustness of predicted and observed food web with respect to the difference in their connectance, been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links. We did this for 12 different food webs from a wide variety of ecosystems. We used three extinction scenarios: random, most connected first, and least connected first.
  4)  We found, as expected, that predicted food webs were more robust than observed food webs, which can be attributed to the higher connectance of the predicted food webs. On average, a one-unit rise in connectance corresponded to a 0.52-unit increase in robustness for the most connected species extinction scenario and a 0.04-unit increase for random species extinction scenarios. Interestingly, the least connected species extinction scenario showed minimal or no effect on increased connectance, on average, for most food webs, except for two.
  5) These results emphasize the significant underestimation of food web robustness caused by undersampling, mitigated by filling in missing links using models. Moreover, differences in the structural properties of the model predicted food webs, such as proportion of basal species and maximum trophic height can have large effects on robustness and therefore could be used to further constrain predicted food webs and perhaps produce more constrained predictions of food web robustness.
